COVID-19 infection and attributable mortality in UK care homes: Cohort study using active surveillance and electronic records (March-June 2020)

### 2021-02-02 manuscript/manuscript.Rmd
first batch corrections from A&A

---
title: "COVID-19 infection and attributable mortality in UK long term care facilities: Cohort study using active surveillance and electronic records (March-June 2020)"
bibliography: references.bib
csl: sage-vancouver.csl
output:
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
  word_document:
    toc: no
    reference_docx: template.docx
link-citations: yes
always_allow_html: yes
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
knit_output_pdf <- F
if(knit_output_pdf){
  options(knitr.table.format = "pdf")
} else {
  options(knitr.table.format = "pandoc")
}
source("../.Rprofile")
devtools::load_all("..")
load_data()
library(dplyr)
library(lubridate)
library(kableExtra)
library(captioner)
library(survival)
library(survminer)
library(ggplot2)
library(finalfit)
# override the finalfit cox PH modelling which defaults to Efron ties
# using Breslow ties instead
# coxphmulti <- function (.data, dependent, explanatory) {
#     requireNamespace("survival")
#     coxph(as.formula(paste0(dependent, "~", paste(explanatory, 
#         collapse = "+"))), data = .data, ties = "breslow")
# }
# assignInNamespace("coxphmulti",coxphmulti,ns="finalfit")

@@

home_first_case <- bind_rows(
  new_cases %>% 
    dplyr::filter(tidyr::replace_na(new_deaths_hospital, 0) > 0 |
                    tidyr::replace_na(new_deaths_home, 0)>0) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(date),
                     source = "CT"),
  residents_overview %>% 
    dplyr::filter(confirmed == 1) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(
      first_case_date = min(covid_first_confirmed, 
                            na.rm = T), 
      source="DAT")
  ) %>% 
  dplyr::group_by(home_code) %>% 
  dplyr::summarise(first_case_date = min(first_case_date))

residents_episodes <- tline %>% 
  filter(rday==1)  %>% 
  left_join(home_first_case) %>%
  mutate(outbreak = dplyr::case_when(
    date >= first_case_date ~ 1L,
    date < first_case_date ~ 0L,
    TRUE ~ 0L
  ))  %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed,
           outbreak) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Uninfected",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(
    covid_status = factor(
      covid_status, 
      levels = c("Uninfected", "Symptomatic (not confirmed)",
                 "Asymptomatic confirmed", "Symptomatic confirmed")),
    ageg = cut_age_denary(compute_age(dob, date_min), 
                          age.min = 75, age.max = 95))   %>% 
  left_join(
    select(reference_homes, home_code, country, region_UK,
           staff_to_bed_ratio, staff_to_bed_ratio_g5, 
           carer_to_bed_ratio, bed_tot_category, 
           beds, pct_occupancy, no_bedrooms, uk_imd_england_quintile,
           occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% ungroup()

@@

## OUTBREAK HOMES 
desc_residents_outbreak_only <- list()
desc_residents_outbreak_only$outbreak_homes <- unique(
  c(home_total_confirmed$home_code[home_total_confirmed$CT_confirmed != 0],
  distinct(filter(new_cases, tidyr::replace_na(new_deaths_hospital, 0) > 0 |
                    tidyr::replace_na(new_deaths_home, 0)>0), home_code)$home_code,
  distinct(filter(residents_overview, confirmed == 1), home_code)$home_code)
)

@@

incidence_CT_residents <- table_nums(
  name = "incidence_CT_residents",
  caption =  paste0("Cumulative incidence and rate of SARS-CoV-2 infections in residents according to FSHCG aggregate data (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
supp_epi_CT_oubreak_only <- table_nums(
  name = "incidence_CT_staff",
  caption =  paste0("Cumulative incidence and rate of SARS-CoV-2 infections among staff according to FSHCG aggregate data (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Characteristics of FSHCG private bed residents by type of LTCF, sex, age, region and status on study exit (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_CFR_age_sex <- table_nums(
  name = "tab_CFR_age_sex",
  caption =  paste0("All-cause case-fatality rates by age and sex among private residents (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  "; ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Risk factors for confirmed infection in private residents: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Risk factors for all-cause mortality in private residents of LTCFs with and without COVID-19 outbreaks: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_episodes$resident_id)),
  ", ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_attributable_deaths <- table_nums(
  name = "tab_attributable_deaths",
  caption =  paste0("Model-based estimates of attributable death in private residents of LTCFs with and without COVID-19 outbreaks (n=", 
  formatbm(dplyr::n_distinct(residents_episodes$resident_id)),
  ", ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))

@@

fig_study_diagram <- fig_nums(
  name = "fig_study_diagram",
  caption = paste0("Study overview: location of FSHCG LTCFs and diagram of data sources")
)

@@

**Background**: Epidemiological data on COVID-19 infection in long-term care facilities (LTCFs) are scarce. We analysed data from a large provider of long-term care to investigate infection and mortality during the first wave of the pandemic.

**Methods**: Cohort study of 179 UK LTCFs with  9,339 residents and 11,604 staff. We used manager-reported daily tallies to estimate the incidence of suspected and confirmed infection and mortality in staff and residents.  Individual-level electronic health records from 8,713 residents were used to model risk factors for confirmed infection, mortality, and estimate attributable mortality.

**Results**: `r formatbm(desc_residents$CT_total_symptomatic)`/`r formatbm(desc_residents$CT_total_residents)` residents developed COVID-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), while `r formatbm(desc_residents$CT_total_confirmed)` residents (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) and `r formatbm(desc_staff$CT_confirmed_total)` staff (`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]) had laboratory-confirmed infections. The rate of incidence of confirmed infection was `r tbrounding(desc_residents$CT_IR_confirmed)` [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`] and `r tbrounding(desc_staff$CT_confirmed_IR)` [`r tbrounding(desc_staff$CT_confirmed_IR_lower)`; `r tbrounding(desc_staff$CT_confirmed_IR_upper)`] per 100,000 person-days in residents and staff respectively. `r length(desc_residents_outbreak_only$outbreak_homes)`/`r desc_residents$total_homes` (`r tbrounding(length(desc_residents_outbreak_only$outbreak_homes)/desc_residents$total_homes*100)`%) LTCFs had at least one COVID-19 infection or COVID-19-related death. Lower staffing ratios and higher occupancy rates were independent risk factors for infection.

`r desc_residents$DAT_CFR$y` of `r desc_residents$DAT_CFR$n` residents with confirmed infection died (case-fatality rate: `r tbrounding(desc_residents$DAT_CFR_rate)`% [`r tbrounding(desc_residents$DAT_CFR_rate_lower)`%; `r tbrounding(desc_residents$DAT_CFR_rate_upper)`%]). Mortality in residents with no direct evidence of infection was 98% higher in care homes with outbreaks versus those without (adjusted HR 2.0 [1.6; 2.4]).

**Conclusions:** Findings suggest many deaths occurred in people who were infected with COVID-19, but not tested. Higher occupancy and lower staffing levels were independently associated with risks of infection. Protecting staff and residents from infection requires regular testing for COVID-19 and fundamental changes to staffing and LTCF occupancy.

@@

Globally the number of COVID-19 cases continues to increase, with substantially higher rates of infection reported in both long-term care facilities (LTCFs) and hospitals.[@PHE2020] In the UK, there are an estimated 400,000 residents living in approximately 11,000 LTCFs for the elderly.[@LaingBuisson2019] Residents of LTCFs are particularly vulnerable to COVID-19 due to older age, high prevalence of comorbidity,[@Zhou2020] and frequent exposure to infection through contact with staff, other residents and contaminated surfaces in the facility. At the peak of the pandemic, deaths recorded in LTCF residents were three times higher than during the preceding year.[@onsDR2020] Staff in LTCFs also had higher aged-standardised rates of COVID-19 related mortality compared to other occupations.[@onsDR2020occ] National statistics suggests two-thirds of excess deaths recorded in residents of LTCFs in the last 6 months involved COVID-19,[@onsDR2020] but this is likely to be an underestimate because many residents were not tested. Understanding the proportion of excess deaths that can be directly and indirectly attributed to COVID-19 infection is important, to fully assess the impact of the pandemic on LTCFs.

Public health strategies to protect the public, residents and staff from COVID-19 include rapid testing, restriction of visitors, and vaccination. Such strategies require knowledge of the burden of and risk factors in residents and staff in LTCFs, linked to outcomes, which may only be drawn from evidence from the pandemic's first wave. Population-based prevalence surveys and studies based on routine data have demonstrated variation in the incidence of infection and case-fatality between countries,[@Fisman2020; @Hollinghurst2020; @Salcher-Konrad2020] but many people with symptoms were not tested, particularly at the start of the pandemic in the context of limited testing capacity. There is no syndromic surveillance for infection in LTCFs in England, and widespread regular testing for SARS-CoV-2 using reverse transcriptase polymerase chain reaction (RT-PCR) was not established for staff and residents in LTCFs until 11 May 2020.[@dhscTesting2020] Prior to this, testing was only available for residents or staff who were admitted to hospital, or as part of Public Health England's (PHE) outbreak investigations which permitted a maximum of five tests per LTCF. Consequently national estimates of incidence and prevalence based on the first wave of infection (February–July 2020) substantially underestimate the burden of infection in residents and staff in LTCFs.

To our knowledge, there are no studies which have employed population-level active surveillance (daily monitoring to identify possible cases of COVID-19 in residents and staff) in LTCFs to investigate the epidemiology and clinical outcomes of both suspected and confirmed COVID-19 infections. We analysed electronic health records from the Four Seasons Health Care Group, one of the UK’s largest independent provider of residential and nursing care, with the aim of identifying strategies to protect staff and residents in LTCFs from future waves of infection. Our objectives were to estimate incidence of and risk factors for infection, and incidence of mortality in the following groups: (A) residents with no evidence of infection; (B) symptomatic residents; (C) asymptomatic residents with confirmed infection; and (D) symptomatic residents with confirmed infection. We also estimated mortality attributable to COVID-19.

@@

Staff and residents living and working in LTCFs for the elderly between `r format(date_min, "%e %B")` and `r format(date_max, "%e %B %Y")`, which were run by the Four Seasons Healthcare Group (FSHCG) were eligible for study inclusion. The FSHCG provides a combination of residential and nursing care (for residents with medical conditions), which is primarily state funded. Most residents are permanent, but a small proportion receive temporary (respite) care.

In 2020, there were 9,568 beds, representing 9% of all beds in England, Scotland and Northern Ireland (supplementary material 1). 90% of FSHCG LTCFs participated in the whole care home testing programme, implying that all staff and residents were tested for COVID-19 at least once between 11 May and 22 June 2020.

@@

FSHCG collects electronic records on all their residents except those occupying beds that are 'block contracted' to the local authority (855 beds). The dataset includes: dates of entry to and exit from the LTCF, sex, date of birth, type of stay (residential/nursing) and care (general, dementia, elderly). Individual-level data on incidents including infections are reported via 'Datix': resident names, LTCF identifier, incident date/time, date of birth, sex, COVID-19 symptoms, test results, resident current location (LTCF/hospital), and death. Individual-level data on residents were linked to Datix reports (supplementary material 1), and used to categorise residents' infection status into four groups:  (A) residents with no evidence of infection (not tested and/or no symptoms); (B) symptomatic residents (symptoms and not tested or tested negative); (C) asymptomatic residents with confirmed infection (no symptoms but tested positive); and (D) symptomatic residents with confirmed infection (symptoms and tested positive) (supplementary material 1). The term 'confirmed' denoted a positive PCR test. Datix was also used to differentiate deaths in hospital from those in the LTCF, and to identify COVID-19 related deaths. 1,492/1,880 (79%) of Datix reports were successfully linked.

@@

On 24 March FSHCG introduced a new reporting system requiring managers of each LTCF to report daily tallies in residents (new symptomatic cases, new confirmed infection in facility, new confirmed infection in hospital, deaths related to COVID-19) and staff (new symptomatic cases, new confirmed cases). The number of occupied beds in each LTCF was reported weekly via the same mechanism. COVID-19 related deaths were defined as death in a resident with confirmed infection or a death attributed to COVID-19 by the coroner. Data on staff deaths were not extracted due to the small number of cases.

Characteristics of each LTCF (number of beds, region, nursing versus residential care) were obtained from FSHCG.

@@

Risk factors considered included individual-level variables (age, sex, general or dementia care, residential versus nursing care) and LTCF characteristics (number of beds, occupancy, bed to staff ratio, UK-wide Index of Multiple Deprivation [@Abel2016]). Baseline LTCF occupancy was computed by averaging weekly occupancy in January-March 2020, before the first COVID-19 case, in order to calculate a ratio of baseline occupancy to the number of bedrooms, and modelled as a continuous variable. We also estimated the ratio of beds to staff as a continuous variable. An outbreak in a LTCF was defined as at least one confirmed infection or COVID-19 related death.

@@

*Infection in staff and residents in LTCFs*

Prevalence, incidence and cumulative incidence were calculated for residents and staff using the aggregate daily tallies. These were the trusted source of information used for national reporting of cases, and encompassed all residents and staff (`r fig_nums("fig_study_diagram", display = "cite")`).[@CQC2020stats] Daily occupancy and numbers of residents at risk of infection were inferred using interpolation and a life table approach, as described in supplementary material 1. Daily occupancy was inferred from the weekly report of bed occupancy using linear interpolation. The total number of residents at risk of infection was unknown, so it was approximated in a multiple-decrement life table (supplementary material 1). The life table allowed us to compute Kaplan-Meier product limit estimators of the cumulative incidence of symptoms, confirmed infections, and COVID-19 related deaths by day based on the aggregate dataset. The incidence rate ratio for LTCF (based on aggregate data) versus community infections was estimated by contrasting the cumulative incidence for confirmed cases in England with estimates from a national household survey for the period `r format(england_results$date_min, "%e %B")`-`r format(england_results$date_max, "%e %B %Y")`.[@cisbulletin2020; @cisdata2020]

Infection incidence was also estimated from the individual-level dataset, but was subject to under-reporting (supplementary material 2). Due to this, individual-level data were only used to estimate age/sex-specific rates of infection and Cox proportional hazards models testing the association with individual and organisational-level risk factors.

>>>>

---
title: "COVID-19 infection and attributable mortality in UK care homes: Cohort study using active surveillance and electronic records (March-June 2020)"
bibliography: references.bib
csl: sage-vancouver.csl
output:
  word_document:
    toc: no
    reference_docx: template.docx
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
link-citations: yes
always_allow_html: yes
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE)
knit_output_pdf <- F
if(knit_output_pdf){
  options(knitr.table.format = "pdf")
} else {
  options(knitr.table.format = "pandoc")
}
source("../.Rprofile")
devtools::load_all()
load_data()
library(dplyr)
library(lubridate)
library(kableExtra)
library(captioner)
library(survival)
library(survminer)
library(ggplot2)
library(finalfit)
# override the finalfit cox PH modelling which defaults to Efron ties
# using Breslow ties instead
# coxphmulti <- function (.data, dependent, explanatory) {
#     requireNamespace("survival")
#     coxph(as.formula(paste0(dependent, "~", paste(explanatory, 
#         collapse = "+"))), data = .data, ties = "breslow")
# }
# assignInNamespace("coxphmulti",coxphmulti,ns="finalfit")

@@

home_first_case <- bind_rows(
  new_cases %>% 
    dplyr::filter(tidyr::replace_na(new_deaths_hospital, 0) > 0 |
                    tidyr::replace_na(new_deaths_home, 0)>0 |
                    tidyr::replace_na(new_confirmed_home, 0)>0 |
                    tidyr::replace_na(new_confirmed_hospital, 0)>0
                  ) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(date),
                     source = "CT"),
  residents_overview %>% 
    dplyr::filter(confirmed == 1) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(
      first_case_date = min(covid_first_confirmed, 
                            na.rm = T), 
      source="DAT")
  ) %>% 
  dplyr::group_by(home_code) %>% 
  dplyr::summarise(first_case_date = min(first_case_date))

home_outbreak_sensitivity <-  new_cases %>% ungroup() %>% 
  arrange(home_code, date) %>% 
  group_by(home_code) %>% 
  mutate(caserunningtot = cumsum(tidyr::replace_na(new_deaths_hospital, 0)+
                    tidyr::replace_na(new_deaths_home, 0))) %>%
  summarise(one_case = min_non_missing(if_else(caserunningtot>0, date, NA_Date_)),
            two_case = min_non_missing(if_else(caserunningtot>1, date, NA_Date_)),
            five_case = min_non_missing(if_else(caserunningtot>4, date, NA_Date_))) %>% 
  left_join(home_first_case)

residents_episodes <- tline %>% 
  filter(rday==1)  %>% 
  left_join(home_outbreak_sensitivity) %>%
  mutate(
  outbreak = dplyr::case_when(
    date >= first_case_date ~ 1L,
    date < first_case_date ~ 0L,
    TRUE ~ 0L
  )
  # outbreak = dplyr::case_when(
  #   date >= two_case ~ 1L,
  #   date < two_case ~ 0L,
  #   TRUE ~ 0L
  # )
  # outbreak = dplyr::case_when(
  #   date >= five_case ~ 1L,
  #   date < five_case ~ 0L,
  #   TRUE ~ 0L)
  ) %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed,
           outbreak) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Uninfected",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(
    covid_status = factor(
      covid_status, 
      levels = c("Uninfected", "Symptomatic (not confirmed)",
                 "Asymptomatic confirmed", "Symptomatic confirmed")),
    ageg = cut_age_denary(compute_age(dob, date_min), 
                          age.min = 75, age.max = 95))   %>% 
  left_join(
    select(reference_homes, home_code, country, region_UK,
           staff_to_bed_ratio, staff_to_bed_ratio_g5, 
           carer_to_bed_ratio, bed_tot_category, 
           beds, pct_occupancy, no_bedrooms, uk_imd_england_quintile,
           occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% ungroup()

@@

## OUTBREAK HOMES 
desc_residents_outbreak_only <- list()
desc_residents_outbreak_only$outbreak_homes <- unique(
  filter(home_outbreak_sensitivity, !is.na(first_case_date) == 1)$home_code
)

@@

incidence_CT_residents <- table_nums(
  name = "incidence_CT_residents",
  caption =  paste0("Cumulative incidence and rate of SARS-CoV-2 infections in residents according to FSHCG aggregate data (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
supp_epi_CT_oubreak_only <- table_nums(
  name = "incidence_CT_staff",
  caption =  paste0("Cumulative incidence and rate of SARS-CoV-2 infections among staff according to FSHCG aggregate data (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Characteristics of FSHCG private bed residents by type of care home, sex, age, region and status on study exit (", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_CFR_age_sex <- table_nums(
  name = "tab_CFR_age_sex",
  caption =  paste0("All-cause case-fatality rates by age and sex among private residents (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  "; ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Risk factors for confirmed infection in private residents: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Risk factors for all-cause mortality in private care home residents with and without COVID-19 outbreaks: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_episodes$resident_id)),
  ", ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))
tab_attributable_deaths <- table_nums(
  name = "tab_attributable_deaths",
  caption =  paste0("Model-based estimates of attributable death in private care home residents with and without COVID-19 outbreaks (n=", 
  formatbm(dplyr::n_distinct(residents_episodes$resident_id)),
  ", ", format(date_min, "%e %b %Y"), "-", format(date_max, "%e %b %Y"), ")"))

@@

fig_study_diagram <- fig_nums(
  name = "fig_study_diagram",
  caption = paste0("Study overview: location of FSHCG care homes and diagram of data sources")
)

@@

**Background**: Epidemiological data on COVID-19 infection in care homes are scarce. We analysed data from a large provider of long-term care for older people to investigate infection and mortality during the first wave of the pandemic.

**Methods**: Cohort study of 179 UK care homes with  9,339 residents and 11,604 staff. We used manager-reported daily tallies to estimate the incidence of suspected and confirmed infection and mortality in staff and residents.  Individual-level electronic health records from 8,713 residents were used to model risk factors for confirmed infection, mortality, and estimate attributable mortality.

**Results**: `r formatbm(desc_residents$CT_total_symptomatic)`/`r formatbm(desc_residents$CT_total_residents)` residents developed COVID-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), while `r formatbm(desc_residents$CT_total_confirmed)` residents (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) and `r formatbm(desc_staff$CT_confirmed_total)` staff (`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]) had laboratory-confirmed infections. The rate of incidence of confirmed infection was `r tbrounding(desc_residents$CT_IR_confirmed)` [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`] and `r tbrounding(desc_staff$CT_confirmed_IR)` [`r tbrounding(desc_staff$CT_confirmed_IR_lower)`; `r tbrounding(desc_staff$CT_confirmed_IR_upper)`] per 100,000 person-days in residents and staff respectively. `r length(desc_residents_outbreak_only$outbreak_homes)`/`r desc_residents$total_homes` (`r tbrounding(length(desc_residents_outbreak_only$outbreak_homes)/desc_residents$total_homes*100)`%) care homes had at least one COVID-19 infection or COVID-19-related death. Lower staffing ratios and higher occupancy rates were independent risk factors for infection.

`r desc_residents$DAT_CFR$y` of `r desc_residents$DAT_CFR$n` residents with confirmed infection died (case-fatality rate: `r tbrounding(desc_residents$DAT_CFR_rate)`% [`r tbrounding(desc_residents$DAT_CFR_rate_lower)`%; `r tbrounding(desc_residents$DAT_CFR_rate_upper)`%]). Mortality in residents with no direct evidence of infection was two-fold higher in care homes with outbreaks versus those without (adjusted HR 2.1 [1.7; 2.5]).

**Conclusions:** Findings suggest many deaths occurred in people who were infected with COVID-19, but not tested. Higher occupancy and lower staffing levels were independently associated with risks of infection. Protecting staff and residents from infection requires regular testing for COVID-19 and fundamental changes to staffing and care home occupancy.

@@

Globally the number of COVID-19 cases continues to increase, with substantially higher rates of infection reported in both care homes and hospitals.[@PHE2020] In the UK, there are an estimated 400,000 residents living in approximately 11,000 care homes for older people with or within on-site nursing care.[@LaingBuisson2019; @Sanford2015] Care home residents are particularly vulnerable to COVID-19 due to older age, high prevalence of comorbidity,[@Zhou2020] and frequent exposure to infection through contact with staff, other residents and contaminated surfaces in the facility. At the peak of the pandemic, deaths recorded in UK care home residents were three times higher than during the preceding year.[@onsDR2020] Staff in UK care homes also had higher aged-standardised rates of COVID-19 related mortality compared to other occupations.[@onsDR2020occ] UK national statistics suggests two-thirds of excess deaths recorded in care home residents in the last 6 months involved COVID-19,[@onsDR2020] but this is likely to be an underestimate because many residents were not tested. Understanding the proportion of excess deaths that can be directly and indirectly attributed to COVID-19 infection is important, to fully assess the impact of the pandemic on care homes.

Public health strategies to protect the public, residents and staff from SARS-CoV-2 include rapid testing, restriction of visitors, and vaccination. Such strategies require knowledge of the burden of and risk factors in residents and staff in care homes, linked to outcomes, which may only be drawn from evidence from the pandemic's first wave. Population-based prevalence surveys and studies based on routine data have demonstrated variation in the incidence of infection and case-fatality between countries,[@Fisman2020; @Hollinghurst2020; @Salcher-Konrad2020] but many people with symptoms were not tested, particularly at the start of the pandemic in the context of limited testing capacity. There is no syndromic surveillance for infection in care homes in England, and widespread regular testing for SARS-CoV-2 using reverse transcriptase polymerase chain reaction (RT-PCR) was not established for staff and residents in care homes until 11 May 2020.[@dhscTesting2020] Prior to this, testing was only available for residents or staff who were admitted to hospital, or as part of Public Health England's (PHE) outbreak investigations which permitted a maximum of five tests per care home. Consequently national estimates of incidence and prevalence based on the first wave of infection (February–July 2020) substantially underestimate the burden of infection in residents and staff in care homes.

To our knowledge, there are no studies which have employed population-level active surveillance (daily monitoring to identify possible cases of COVID-19 in residents and staff) in care homes to investigate the epidemiology and clinical outcomes of both suspected and confirmed COVID-19 infections. We analysed electronic health records from the Four Seasons Health Care Group, one of the UK’s largest for-profit provider of residential and nursing care, with the aim of identifying strategies to protect staff and residents in care homes from future waves of infection. Our objectives were to estimate incidence of and risk factors for infection, and incidence of mortality in the following groups: (A) residents with no evidence of infection; (B) symptomatic residents; (C) asymptomatic residents with confirmed infection; and (D) symptomatic residents with confirmed infection. We also estimated mortality attributable to COVID-19.

@@

Staff and residents living/working in care homes for older people run by the Four Seasons Healthcare Group (FSHCG) between `r format(date_min, "%e %B")` and `r format(date_max, "%e %B %Y")` were eligible for study inclusion. The FSHCG provides a combination of residential and nursing care (for residents with medical conditions), which is predominantly state-funded. Most residents are permanent, but a small proportion receive temporary (respite) care.

In 2020, there were 9,568 beds, representing 9% of all registered care home beds in England, Scotland and Northern Ireland (supplementary methods). 90% of FSHCG care homes participated in the whole care home testing programme, implying that all staff and residents were tested for COVID-19 at least once between 11 May and 22 June 2020.

@@

FSHCG collects electronic records on residents occupying 'private' beds, excluding those occupying beds that are 'block contracted' to the local authority (855 beds, see Figure 1). Records includes: dates of entry to and exit from the care home, sex, date of birth, type of stay (residential/nursing) and care (general, dementia, older residents). Individual-level data on incidents including infections are reported via 'Datix': resident names, care home identifier, incident date/time, date of birth, sex, COVID-19 symptoms (9 multiple choices), test results, resident current location (care home/hospital), and death. Individual-level data on residents were linked to Datix reports (supplementary methods), and used to categorise residents' infection status into four groups: (A) residents with no evidence of infection (not tested and/or no symptoms); (B) symptomatic residents (symptoms and not tested or tested negative); (C) asymptomatic residents with confirmed infection (no symptoms but tested positive); and (D) symptomatic residents with confirmed infection (symptoms and tested positive) (supplementary methods). The term 'confirmed' denoted a positive PCR test. Datix was also used to differentiate deaths in hospital from those in the care home, and to identify COVID-19 related deaths. 1,492/1,880 (79%) of Datix reports were successfully linked.

@@

On 24 March FSHCG introduced a new reporting system requiring managers of each care home to report daily tallies in residents (new symptomatic cases, new confirmed infection in facility, new confirmed infection in hospital, deaths related to COVID-19) and staff (new symptomatic cases, new confirmed cases). The number of occupied beds in each care home was reported weekly via the same mechanism. COVID-19 related deaths were defined as death in a resident with confirmed infection or a death attributed to COVID-19 by the coroner. Data on staff deaths were not extracted due to the small number of cases.

Characteristics of each care home (number of beds, region, nursing versus residential care) were obtained from FSHCG.

@@

Risk factors considered included individual-level variables (age, sex, general or dementia care, residential versus nursing care) and care home characteristics (number of beds, occupancy, bed-to-staff ratio, UK-wide Index of Multiple Deprivation [@Abel2016]). Baseline care home occupancy was computed by averaging weekly occupancy in January-March 2020, before the first COVID-19 case, in order to calculate a ratio of baseline occupancy to the number of bedrooms, and modelled as a continuous variable. We also estimated the ratio of beds to staff as a continuous variable.

To adjust for case under-ascertainment, we created a dummy variable indicating 'affected care homes', which we define throughout the manuscript as care homes that recorded $\geq$1 confirmed infection or COVID-19 related death during the period.

@@

*Infection in staff and residents in care homes*

Prevalence, incidence and cumulative incidence were calculated for residents and staff using the aggregate daily tallies. These were the trusted source of information used for national reporting of cases, and encompassed all residents and staff (`r fig_nums("fig_study_diagram", display = "cite")`).[@CQC2020stats] Daily occupancy and numbers of residents at risk of infection were inferred using interpolation and a life table approach, as described in supplementary methods. Daily occupancy was inferred from the weekly report of bed occupancy using linear interpolation. The total number of residents at risk of infection was unknown, so it was approximated in a multiple-decrement life table (supplementary methods). The life table allowed us to compute Kaplan-Meier product limit estimators of the cumulative incidence of symptoms, confirmed infections, and COVID-19 related deaths by day based on the aggregate dataset. The incidence rate ratio for care home (based on aggregate data) versus community infections was estimated by contrasting the cumulative incidence for confirmed cases in England with estimates from a national household survey for the period `r format(england_results$date_min, "%e %B")`-`r format(england_results$date_max, "%e %B %Y")`.[@cisbulletin2020; @cisdata2020]

Infection incidence was also estimated from the individual-level dataset, but was subject to under-reporting (Table S1b). Due to this, individual-level data were only used to estimate age/sex-specific rates of infection and Cox proportional hazards models testing the association with individual and organisational-level risk factors.
